---
title: "Example Notebook"
author: "Your Name Here"
date: "July 28, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(ggplot2)
```

# Overview

What is the point of this research? What questions are you trying to answer?

## Hypothesis

What are you proposing?

## Methodology

How did you gather the data? How did you go about approaching the problem?

# Data

Load your data here. Where did it come from? Did you do anything special to it?

```{r load_data}
meetups <- read.csv("tech_meetups/meetup_groups.csv")
```

Validate your data
```{r validate_data}

paste("Number of rows:", nrow(meetups))

# Columns
paste(names(meetups), sep=" ")
```

Summary is a quick way to make sure it's what you expect.

What states are there?
```{r}
summary(meetups$state)
```

Anything non-US?
```{r}
summary(meetups$country)
```

Anything non-tech?
```{r}
summary(meetups$category.name)
```

# Analysis

How did you go about exploring your question?

How many meetup groups are in each state?

```{r}
ggplot(meetups, aes(x=state, fill=status)) +
  geom_bar() +
  xlab("State") +
  guides(fill=guide_legend(title="Status"))
```

How many members are in each state?

```{r}
ggplot(meetups, aes(x=state, y=members, fill=is.na(next_event.id))) +
  geom_bar(stat="identity") +
  xlab("State") +
  ylab("Members")
  guides(fill=guide_legend(title="Has Next Event?"))
```

Wow, California has a lot!

```{r}
meetups %>% filter(state=="CA") %>% summarise(sum(members))
```

What is the age of the meetups?

```{r}
# convert from epoch time
meetups <- meetups %>%
  mutate(created_date = as.POSIXct(created/1000, origin="1970-01-01"),
         created_y = format(created_date, "%Y"),
         created_my = format(created_date, "%b-%Y"))
```


```{r}
ggplot(meetups, aes(x=created_y, fill=state)) +
  geom_bar() +
  xlab("Created") +
  ylab("Meetup Groups")
  guides(fill=guide_legend(title="State"))
  
```


```{r}
ggplot(meetups %>% filter(state=="CA"), aes(city, members)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

## Your TURN!

Make this work with the Meetup data:

Plot a histogram with one variable:

```{r}
ggplot(my_data, aes(x=column_name_for_thing_to_count) +
  geom_bar() +
  xlab("Thing I'm Counting")
```

Group and Summarize something in the dataset:

```{r}
# things per month, creates a summarized version of your dataset

things_per_month_summary <- my_things %>%
  group_by(month, person) %>% # how do you want to group?
  summarise(num_things = n(), # number of rows total
			num_unique_things = n_distinct(things), # number of unique things (doesn't count repeats)
            biggest_thing=max(thing),
			smallest_thing=min(thing),
			average_thing=mean(thing),
			middle_thing=med(thing))

# If you don't want to summarize but want to keep rows intact, you can use "mutate" instead
things_per_month_all <- my_things %>%
  group_by(month, person) %>% # how do you want to group?
  mutate(num_things = n(), # number of rows total in the grouping
			num_unique_things = n_distinct(thing), # number of unique things (doesn't count repeats)
            biggest_thing=max(thing),
			smallest_thing=min(thing),
			average_thing=mean(thing),
			middle_thing=med(thing))
```

Plot a summary using a bar chart:

```{r}
# to use a bar plot with summaries, you need to use the keyword "identity"

ggplot(my_data_summary, aes(x=numbers, y=thing) +
  geom_bar(stat="identity") +
  xlab("Thing I'm Numbering") +
  ylab("Numbers")
	   
# to plot a grouped value, use fill
# use position="dodge" for bars next to each other, 
# position="stack" stacks them on top
ggplot(my_data_summary, aes(x=numbers, y=thing, fill=thing_type) +
  geom_bar(stat="identity", position="stack") +
  xlab("Thing I'm Numbering") +
  ylab("Numbers") 

# if your things are in the wrong order you can use reorder()
# a "-" in front means descending, otherwise it's ascending (smallest first)
ggplot(my_data_summary, aes(x=numbers, y=thing, fill=reorder(thing_type, -num_things)) +
  geom_bar(stat="identity", position="stack") +
  xlab("Thing I'm Numbering") +
  ylab("Numbers")  
```


Plot 2 different columns using a line graph:

```{r}
ggplot(data = my_data, aes(x = x_axis_thing) +
  geom_point(aes(y = numerical_value1, color="Thing 1"), group=1) +
  geom_line(aes(y = numerical_value1, color="Thing 1"), group=1) +
  geom_point(aes(y = numerical_value2, color="Thing 2"), group=2) +
  geom_line(aes(y = numerical_value2, color="Thing 2"), group=2) +
  ylab("Numbers") +
  xlab("Thing I'm Numbering")
```



